---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
### Load Packages and Libraries
```{r , echo=TRUE}
library(ggplot2)
```
### Loading and preprocessing the data

### 1.Reading File

```{r , echo=TRUE}
activity_data<-read.csv("activity.csv", header=TRUE)
head(activity_data)

```


### What is mean total number of steps taken per day?

### 2. Histogram of the total number of steps taken each day

```{r , echo=TRUE}

per_day_steps<- aggregate(steps ~ date, activity_data, sum, na.rm=TRUE)
ggplot(per_day_steps, aes(x=steps)) + geom_histogram(binwidth=500,color="black", fill="white")

```

### 3. Mean and median of number of steps taken each day
```{r , echo=TRUE}

mean_total<-mean(per_day_steps$steps, na.rm=TRUE)
median_total<-median(per_day_steps$steps, na.rm=TRUE)


```

### Mean of number of steps taken each day :
```{r , echo=TRUE}
round(mean_total)

```

### Median of number of steps taken each day :
```{r , echo=TRUE}
round(median_total)

```

### What is the average daily activity pattern?


### Time series plot of the average number of steps taken in 5-minutes time interval
```{r , echo=TRUE}

active_activity_data<-activity_data[complete.cases(activity_data),]

avg5min_steps<-aggregate(steps~interval, active_activity_data, mean)

ggplot(avg5min_steps, aes(interval, steps, type = "l")) + 
    geom_line(color="red") + 
    xlab("5-Minutes Time Interval") +
    ylab("Average Number Of Steps") +
    ggtitle("Time series plot of the average number of steps taken in 5-minutes time interval")


```


### 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r , echo=TRUE}

max_steps_interval<-which.max(avg5min_steps$steps)
max_steps_interval


print (paste("Interval with the highest average number of steps is ", avg5min_steps[max_steps_interval, ]$interval, " and the no of steps for that interval is ", round(avg5min_steps[max_steps_interval, ]$steps, digits = 0)))
```
### Imputing missing values

### 6. Code to describe and show a strategy for imputing missing data
```{r , echo=TRUE}
#Find Missing Values
missing_values_data<-activity_data[!complete.cases(activity_data),]
print(paste("Total rows with missing value are :", nrow(missing_values_data)))

#Find rows with NA and imputing missing values with the mean 
impute_activity_data<-activity_data
impute_activity_data$steps[is.na(impute_activity_data$steps)]<-mean(impute_activity_data$steps, na.rm=TRUE)

#Check if there is any NA value left

impute_activity_data$steps <- as.numeric(impute_activity_data$steps)
impute_activity_data$interval <- as.numeric(impute_activity_data$interval)
colSums(is.na(impute_activity_data))

# New data with filled values
summary(impute_activity_data)
```
### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r , echo=TRUE}

impute_steps_daily<-aggregate(steps~date, impute_activity_data, sum)
ggplot(impute_steps_daily, aes(x=steps)) + geom_histogram(color="red", fill="white") +
  ggtitle("Total Number of Steps Each Day After Imputing Values")+
  xlab("Steps")+
  ylab("Count")

```

### Are there differences in activity patterns between weekdays and weekends?

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r , echo=TRUE}
# create a factor variable having 2 labels( Weekday and Weekend)
library(timeDate)
impute_activity_data$day <- ifelse (isWeekday(impute_activity_data$date),"Weekday", "Weekend")
plot_data<-aggregate(steps~interval+day, impute_activity_data, mean)


# PLOT

ggplot(plot_data, aes(interval, steps, color=day), type="l")+
  geom_line()+
  facet_grid(day~.)+ 
  xlab("5 Minutes Interval ")+
  ylab("Steps Taken")+
  ggtitle("Comparison of average number of steps/per 5-minute interval weekdays Vs weekends")
  

```

